Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2022-05-04 article/draft.tex
ref added

Furthermore, mainstream platforms are increasingly engaging in editorial-like tasks by implementing targeted policies to insure that each platform's rules are not violated. Community guidelines of Facebook, Twitter and YouTube can be summarized in a handful of categories, regarding safety, privacy and authenticity; which include sub-categories such as violence, terrorism, child sexual exploitation, abuse, harassment, hateful conduct, suicide or self-harm, illegal or regulated goods and services, platform manipulation and spam (see Appendix \ref{links} for references). While specific to each platform, the previously cited categories correspond in most cases to well defined concepts that fall into legal frameworks in many countries, unlike misinformation. The intricacies of constructing a legal framework for misinformation arises from the difficulty of identifying and qualifying a piece of online content as false or misleading, among an overwhelming quantity of daily produced content, without infringing existing laws.\footnote{See \href{https://www.poynter.org/ifcn/anti-misinformation-actions/}{A guide to anti-misinformation actions around the world} on the website of Poynter Institute.} In particular, a number of recent studies point towards the idea that misinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense.\footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.'' Furthermore, the European commission recognizes the spread of online disinformation as a problem and has put together in 2018 a \href{https://digital-strategy.ec.europa.eu/en/policies/code-practice-disinformation}{Code of practice on Disinformation}, which is a set of self-regulatory standards to fight disinformation.}

@@

We are aware that full transparency regarding operational aspects of misinformation related interventions can backfire.
This is because the aimed users might find ways to circumvent the interventions. 
We noticed that some YouTube channels are trying to avoid the likely automated COVID-19 information panels, by using notational variants such as ``C.O.V.I.D'' or ``C O V I D''. 
Similarly, we came across notational variants on Twitter of the domain name {\it off-guardian.org}
\footnote{Here is an example of a failed fact-check of one of the articles of the website {\it off-guardian.org} : \href{https://www.politifact.com/factchecks/2020/jul/07/blog-posting/covid-19-tests-are-not-scientifically-meaningless/}{politifact.com/factchecks/2020/jul/07/blog-posting/covid-19-tests-are-not-scientifically-meaningless/}}
such as {\it off-guardian. org}, {\it off-guardian .org}, {\it off-guardian./org}, {\it off-guardian*org}.\footnote{For this case, we failed to understand precisely what kind of intervention Twitter users were circumventing, since they could share in a Tweet an article redirecting to the website {\it off-guardian.org} with the correct notation and only a warning message would appear when one clicks on it.}
Hence, a balance is needed between: $(i)$ fully transparent communication about misinformation related interventions so that the users can understand how the platforms are regulating their content, and $(ii)$ not explicitly giving too much information which would allow the policies to be bypassed.

@@

\end{landscape}
%https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works
\noindent {\bf Funding} \\
This research was supported by the Make Our Planet Great Again’ French state aid managed by the Agence Nationale de la Recherche under the ’Investissements d’avenir’ program with the reference ANR-19-MPGA-0005. \bigskip

@@

\noindent {\bf Ethics} \\
The data collection and processing complied with the EU General Data Protection Regulation (GDPR).
\end{document}

>>>>

Furthermore, mainstream platforms are increasingly engaging in editorial-like tasks by implementing targeted policies to insure that each platform's rules are not violated. Community guidelines of Facebook, Twitter and YouTube can be summarized in a handful of categories, regarding safety, privacy and authenticity; which include sub-categories such as violence, terrorism, child sexual exploitation, abuse, harassment, hateful conduct, suicide or self-harm, illegal or regulated goods and services, platform manipulation and spam (see Appendix \ref{links} for references). While specific to each platform, the previously cited categories correspond in most cases to well defined concepts that fall into legal frameworks in many countries, unlike misinformation (see Fathaigh et al. (2021)~\cite{fathaigh} for a discussion for the case of the EU). The intricacies of constructing a legal framework for misinformation arises from the difficulty of identifying and qualifying a piece of online content as false or misleading, among an overwhelming quantity of daily produced content, without infringing existing laws.\footnote{See \href{https://www.poynter.org/ifcn/anti-misinformation-actions/}{A guide to anti-misinformation actions around the world} on the website of Poynter Institute.} In particular, a number of recent studies point towards the idea that misinformation is a small subset of the total supply of information on online social networking platforms (e.g. Grinberg et al. (2019)~\cite{grinberg} and Broniatowski et al. (2020)~\cite{broniatowski}). Yet, this seemingly small subset is generating great concern in traditional media and in society in a broader sense.\footnote{For example see the February 2020 \href{https://www.who.int/director-general/speeches/detail/munich-security-conference}{speech} of the Director General of the WHO at the Munich Security Conference, where he says ``But we’re not just fighting an epidemic; we’re fighting an infodemic.'' Furthermore, the European commission recognizes the spread of online disinformation as a problem and has put together in 2018 a \href{https://digital-strategy.ec.europa.eu/en/policies/code-practice-disinformation}{Code of practice on Disinformation}, which is a set of self-regulatory standards to fight disinformation.}

@@

We are aware that full transparency regarding operational aspects of misinformation related interventions can backfire.
This is because the aimed users might find ways to circumvent the interventions. 
We noticed that some YouTube channels are trying to avoid the likely automated COVID-19 information panels, by using notational variants such as ``C.O.V.I.D'' or ``C O V I D''. 
Similarly, we came across notational variants on Twitter of the domain name {\it off-guardian.org}
\footnote{Here is an example of a failed fact-check of one of the articles of the website {\it off-guardian.org} : \href{https://www.politifact.com/factchecks/2020/jul/07/blog-posting/covid-19-tests-are-not-scientifically-meaningless/}{politifact.com/factchecks/2020/jul/07/blog-posting/covid-19-tests-are-not-scientifically-meaningless/}}
such as {\it off-guardian. org}, {\it off-guardian .org}, {\it off-guardian./org}, {\it off-guardian*org}.\footnote{For this case, we failed to understand precisely what kind of intervention Twitter users were circumventing, since they could share in a Tweet an article redirecting to the website {\it off-guardian.org} with the correct notation and only a warning message would appear when one clicks on it.} Furthermore there is little information about how algorithms are practically used as a tool for content moderation; since they can be used to reduce the visibility of some content or even prevent problematic content from being recommended.  Hence researchers have to resort to proxy measures and simulated experiments to have a better understanding the impact of algorithmic design. Based on a sample of $23$ YouTube studies, Yesilada and Lewandowsky (2022)~\cite{yesilada} find that YouTube recommender system could lead users to problematic content, but highlight the limitations of this conclusion since there is an incomplete understanding of the recommendation system and models built might not reflect the actual functioning of YouTube recommendation algorithm. 
%point here about algorithms and cite Yesilada and Lewandowsky (2022)~\cite{yesilada}
Hence, a balance is needed between: $(i)$ fully transparent communication about misinformation related interventions so that both users and researchers can understand and investigate how the platforms are regulating content, and $(ii)$ not explicitly giving too much information which would allow the policies to be bypassed.

@@

\end{landscape}
%https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works
\noindent {\bf Acknowledgments} \\
We thank our colleagues at Sciences Po m\'{e}dialab for their feedback and suggestions. \bigskip

\noindent {\bf Funding} \\
This research was supported by the Make Our Planet Great Again’ French state aid managed by the Agence Nationale de la Recherche under the ’Investissements d’avenir’ program with the reference ANR-19-MPGA-0005. \bigskip

@@

\noindent {\bf Ethics} \\
The data collection and processing complied with the EU General Data Protection Regulation (GDPR).

\end{document}
